Predicting the Quality of CDRH3 Antibody Loop Structural Models

### 2022-10-05 paper.tex
A4 and CDR-H3 standardization

\documentclass[12pt]{article}
\bibliographystyle{unsrt}
\usepackage{graphicx}
\usepackage{url}
%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}

@@

\begin{abstract}
  Therapeutic antibodies have shown an unprecedented pace of
  development and have brought new hope for the treatment of numerous
  diseases. The bioinformatic tools for modelling antibody structures
  have become invaluable for antibody engineering and the development
  of therapeutic antibodies. The antigen-binding site consists of six
  hypervariable loops, also known as the Complementary Determining
  Regions (CDR), all of which can be modelled with adequate accuracy,
  except for one. It remains markedly difficult to model the third CDR
  loop of the antibody heavy chain. The CDRH3 differs in length, has
  far greater sequence variability and has such a great structural
  diversity that modelling it is considerably harder.

  Many sophisticated approaches for antibody modelling, such as the
  abYmod software, have been developed. Although such efforts have
  improved prediction accuracy the results for the CDRH3 loop are
  still inconsistent and require further improvement. Providing a
  confidence score for the structure predictions would aid in
  differentiating well-modelled structures from incorrectly modelled
  structures, giving the abYmod user a clearer understanding of the
  generated model reliability.

  We present a model quality predictor, combining domain knowledge
  with machine learning techniques to predict the accuracy of CDRH3
  models generated by antibody modelling software such as abYmod. The
  newly developed predictor scored a Mathews Correlation Coefficient
  of 0.99, and can thus be described as highly reliable. The predictor
  is made available at \url{http://www.bioinf.org.uk/abs/qualiloop/}
\end{abstract}

@@

Antibodies consist of a heavy and a light chain, which are linked by
disulphide bonds. The N-terminal domain of each makes up the variable
fragment (Fv), which contains the complementarity determining regions
(CDRs). The antigen binding site is composed of six CDRs, also known
as hypervariable loops. All except one of these loops can be clustered
into a limited number of canonical structures. Therefore, modelling
these loops with adequate accuracy is commonly achievable
\cite{North2011},\cite{Weitzner2015}. However, the CDR loop 3 of the
heavy chain (CDRH3) has a far greater sequence variability due to the
processes of V(D)J recombination and somatic hyper‐mutation and its
structure has remained unclassifiable \cite{Finn2016}. The variety in
structure is so great, that its structural diversity is remarkable
even compared to other protein loops \cite{Regep2017}. It was found
that over 75\% of CDRH3 loops do not have a sub-Angstrom non-antibody
structural neighbour, as well as that 30\% of CDRH3 loops have a
completely unique structure compared with under 3\% for all other
loops on average \cite{Regep2017}.

Apart from being the most structurally diverse, the H3 loop is also
the most important for antigen binding, being located at the center of
the binding site and forming the most contacts to the antigen
\cite{MacCallum1996}. In fact, it was demonstrated that differences in
this loop alone were sufficient to enable otherwise identical
antibodies to distinguish between various antigens \cite{Xu2000}.

According to the Kabat definition, the CDRH3 loop is made up of the
residues 95-105 (Kabat numbering scheme\cite{Kabat1992}) in the heavy
chain, with a potential insertion site at position 100. The
possibility of such an insertion of a varying number of residues leads
to a large range of loop lengths, with bovine antibodies being
exceptionally long \ref{fig:loopdist}.

\begin{figure}
  \includegraphics[scale=1]{loopdist.eps}
  \centering
  \caption { Distribution of CDR-H3 loop lengths in mouse (left),
    human (centre) and cow (right). Mouse and human antibody CDRH3s
    have a unimodal, yet almost normal distribution with a range of
    ca. 4-28 and 4-38 respectively. The length cut-off for CDRH3s
    depicted above is 30 amino acids. Bovine antibodies with
    significantly longer CDR-H3 loops than depicted here exist,
    reaching a length of 67 amino acids and
    above. \protect\cite{Wong2015}}  
  \label{fig:loopdist}
\end{figure}

For shorter loops, a higher prediction accuracy can be achieved than
for longer CDRH3 loops. This was also shown by the Antibody Modelling
Assessments (AMA), two blind contests that required researchers to
build structural models from antibody sequences. The CDR-H3 loop
modelling quality achieved at the contests was on average much lower
for loops of longer lengths \cite{Almagro2011},\cite{Almagro2014}.

@@

In this project abYmod is used, which is to be found at
http://abymod.abysis.org/. The program utilizes extensive canonical
class definitions, VH/VL angle prediction and a large database of loop
structures (LoopDB) for CDR-H3 modelling to achieve optimal
results. Upon inputting an antibody sequence, abYmod assigns the
canonical class using a set of key residues \cite{Martin1996}and where
an exact match is not possible, a nearest match is made. Then, the
program identifies the 10 best overall-matching PDB files according to
sequence identity for the light and heavy chain. Of these, the best
template is then identified for each CDR using sequence similarity and
identity. In general, modelling the antibody using the single best
overall-matching template works best and the CDR-specific templates
are used if there is no canonical match. The VH/VL packing angle is
then determined either using machine learning or the chosen template
structure for one of the chains. CDR specific templates, if selected,
are grafted onto the framework. If there is no template of the correct
length for CDR-H3, the loop is built using the LoopDB database,
containing CDRH3-like loops from all proteins. Finally, Gromacs energy
minimization software is used to optimize the model. This method has
proven very effective and preliminary analysis suggests the method
achieves comparable results or outperforms other modelling software
(see results section).

@@

Modelling the H3 loop is a hurdle for in silico development of
therapeutic antibodies. Currently, there is no definite, reliable way
to determine how accurate a generated structural model is within the
H3 region. Therefore, the aim of this project will be to produce a
user-friendly predictor of H3 model quality. The predictor will give
the user an RMSD-range in {\AA}ngstr\"{o}ms, in which the generated model lies
with a high probability. Making such a confidence score available via
the web interface of the in-house modelling software abYmod is a
future goal. Such a score is not provided by most modelling programs
and would thus be a novel addition.  This information can guide the
user in the antibody engineering process. The user has the choice to
determine whether the model is to be used in the intended way, or
whether the model should be re-worked.

@@

\begin{figure}
  \includegraphics[scale=0.9]{logo.eps}
  \caption {Sequence Logo of the CDRH3 loop sequence. Data on amino
    acid occurrence taken from http://abymod.abysis.org/. Visualized
    using Seq2Logo.}
  \label{fig:logo}
\end{figure}

>>>>

\documentclass[12pt]{article}
\bibliographystyle{unsrt}
\usepackage{a4}
\usepackage{graphicx}
\usepackage{url}
%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}

@@

\begin{abstract}
  Therapeutic antibodies have shown an unprecedented pace of
  development and have brought new hope for the treatment of numerous
  diseases. The bioinformatic tools for modelling antibody structures
  have become invaluable for antibody engineering and the development
  of therapeutic antibodies. The antigen-binding site consists of six
  hypervariable loops, also known as the Complementary Determining
  Regions (CDR), all of which can be modelled with adequate accuracy,
  except for one. It remains markedly difficult to model the third CDR
  loop of the antibody heavy chain. The CDR-H3 differs in length, has
  far greater sequence variability and has such a great structural
  diversity that modelling it is considerably harder.

  Many sophisticated approaches for antibody modelling, such as the
  abYmod software, have been developed. Although such efforts have
  improved prediction accuracy the results for the CDR-H3 loop are
  still inconsistent and require further improvement. Providing a
  confidence score for the structure predictions would aid in
  differentiating well-modelled structures from incorrectly modelled
  structures, giving the abYmod user a clearer understanding of the
  generated model reliability.

  We present a model quality predictor, combining domain knowledge
  with machine learning techniques to predict the accuracy of CDR-H3
  models generated by antibody modelling software such as abYmod. The
  newly developed predictor scored a Mathews Correlation Coefficient
  of 0.99, and can thus be described as highly reliable. The predictor
  is made available at \url{http://www.bioinf.org.uk/abs/qualiloop/}
\end{abstract}

@@

Antibodies consist of a heavy and a light chain, which are linked by
disulphide bonds. The N-terminal domain of each makes up the variable
fragment (Fv), which contains the complementarity determining regions
(CDRs). The antigen binding site is composed of six CDRs, also known
as hypervariable loops. All except one of these loops can be clustered
into a limited number of canonical structures. Therefore, modelling
these loops with adequate accuracy is commonly achievable
\cite{North2011},\cite{Weitzner2015}. However, the CDR loop 3 of the
heavy chain (CDR-H3) has a far greater sequence variability due to the
processes of V(D)J recombination and somatic hyper‐mutation and its
structure has remained unclassifiable \cite{Finn2016}. The variety in
structure is so great, that its structural diversity is remarkable
even compared to other protein loops \cite{Regep2017}. It was found
that over 75\% of CDR-H3 loops do not have a sub-Angstrom non-antibody
structural neighbour, as well as that 30\% of CDR-H3 loops have a
completely unique structure compared with under 3\% for all other
loops on average \cite{Regep2017}.

Apart from being the most structurally diverse, the CDR-H3 loop is also
the most important for antigen binding, being located at the center of
the binding site and forming the most contacts to the antigen
\cite{MacCallum1996}. In fact, it was demonstrated that differences in
this loop alone were sufficient to enable otherwise identical
antibodies to distinguish between various antigens \cite{Xu2000}.

According to the Kabat definition, the CDR-H3 loop is made up of the
residues 95-105 (Kabat numbering scheme\cite{Kabat1992}) in the heavy
chain, with a potential insertion site at position 100. The
possibility of such an insertion of a varying number of residues leads
to a large range of loop lengths, with bovine antibodies being
exceptionally long \ref{fig:loopdist}.

\begin{figure}
  \includegraphics[scale=1]{loopdist.eps}
  \centering
  \caption { Distribution of CDR-H3 loop lengths in mouse (left),
    human (centre) and cow (right). Mouse and human antibody CDR-H3s
    have a unimodal, yet almost normal distribution with a range of
    ca. 4-28 and 4-38 respectively. The length cut-off for CDR-H3s
    depicted above is 30 amino acids. Bovine antibodies with
    significantly longer CDR-H3 loops than depicted here exist,
    reaching a length of 67 amino acids and
    above. \protect\cite{Wong2015}}  
  \label{fig:loopdist}
\end{figure}

For shorter loops, a higher prediction accuracy can be achieved than
for longer CDR-H3 loops. This was also shown by the Antibody Modelling
Assessments (AMA), two blind contests that required researchers to
build structural models from antibody sequences. The CDR-H3 loop
modelling quality achieved at the contests was on average much lower
for loops of longer lengths \cite{Almagro2011},\cite{Almagro2014}.

@@

In this project abYmod is used, which is to be found at
http://abymod.abysis.org/. The program utilizes extensive canonical
class definitions, VH/VL angle prediction and a large database of loop
structures (LoopDB) for CDR-H3 modelling to achieve optimal
results. Upon inputting an antibody sequence, abYmod assigns the
canonical class using a set of key residues \cite{Martin1996}and where
an exact match is not possible, a nearest match is made. Then, the
program identifies the 10 best overall-matching PDB files according to
sequence identity for the light and heavy chain. Of these, the best
template is then identified for each CDR using sequence similarity and
identity. In general, modelling the antibody using the single best
overall-matching template works best and the CDR-specific templates
are used if there is no canonical match. The VH/VL packing angle is
then determined either using machine learning or the chosen template
structure for one of the chains. CDR specific templates, if selected,
are grafted onto the framework. If there is no template of the correct
length for CDR-H3, the loop is built using the LoopDB database,
containing CDR-H3-like loops from all proteins. Finally, Gromacs energy
minimization software is used to optimize the model. This method has
proven very effective and preliminary analysis suggests the method
achieves comparable results or outperforms other modelling software
(see results section).

@@

Modelling the CDR-H3 loop is a hurdle for in silico development of
therapeutic antibodies. Currently, there is no definite, reliable way
to determine how accurate a generated structural model is within the
H3 region. Therefore, the aim of this project will be to produce a
user-friendly predictor of CDR-H3 model quality. The predictor will give
the user an RMSD-range in {\AA}ngstr\"{o}ms, in which the generated model lies
with a high probability. Making such a confidence score available via
the web interface of the in-house modelling software abYmod is a
future goal. Such a score is not provided by most modelling programs
and would thus be a novel addition.  This information can guide the
user in the antibody engineering process. The user has the choice to
determine whether the model is to be used in the intended way, or
whether the model should be re-worked.

@@

\begin{figure}
  \includegraphics[scale=0.9]{logo.eps}
  \caption {Sequence Logo of the CDR-H3 loop sequence. Data on amino
    acid occurrence taken from http://abymod.abysis.org/. Visualized
    using Seq2Logo.}
  \label{fig:logo}
\end{figure}
